---
title: "ESM204_HW4"
author: "Duerksen-Salm, Massion, Shoemaker"
date: "5/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Attach packages

library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(effsize)
library(kableExtra)
library(ggbeeswarm)
library(ggridges)
library(patchwork)
library(car)
library(corrplot)
library(stargazer)
library(sf)
library(gganimate)
library(transformr)

# SOMETHING
```

```{r}
# Temp function

temp = function(t,T)
{
  out = min(t*T/100,T)
  return(out)
}

# Test temp function

temp(t = 50, T = 5)

# It works! Huzzah! 

```

```{r}
# Keep function 

keep = function(temp)
{
  out = exp(-0.05*temp^2)
  return(out)
}
 
# Test keep function 

keep(temp = 10) 

# I got a number? Same as I got in wolfram alpha

```

```{r}
# Consumption function

consump = function(keep,t)
{
  out = keep*exp(0.01*t)
  return(out)
}

# Test consump function

consump(keep = 10, t = 5)

# I think it works (cross reference with WA). Seems suspicious that I haven't had any issues 
```

```{r}
# Utility function 

util = function(C)
{
  out = ((C^(1-0.5))/(1-0.5))
  return(out)
}

# Test util function 

util(C = 5)

# It works? 
```

```{r}
# PV of util (ramsey) - we are given all the numbers here

delta = 0.005
eta = 0.5
g = 0.01
r = delta + (eta*g)
```


1. PLOTS

(a) Plot temperature over time for no climate change (T = 0), with modest climate change (T = 2), and with extreme climate change (T = 8).
```{r}
no_CC <- temp(t = 100, T = 0)
# No Climate Change :: tao(t) = O


modest_CC <- temp(t = 100, T = 2)
# Modeterate Climate Change :: tao(t) = 2
temp(t = 25, T = 2)
# t = 25 -> temp = 0.5
# (100,2) // (25,0.5)
slope_modest_cc <- (2-0.5)/(100-25)
#slope = 0.02
#y-int = 0


extreme_CC <- temp(t = 100, T = 8)
# Extreme Climate Change :: tao(t) = 8
temp(t = 25, T = 8)
# t = 25 -> temp = 2
# (100,8) // (25,2)
slope_extreme_cc <- (8-2)/(100-25)
# slope = 0.08
# y-int = 0


### PLUTO ###

# No climate change:
T.1 = 0
# Modest change:
T.2 = 2
# Extreme change:
T.3 = 8

t <- seq(1,200, by = 1)
# No climate change:
Temp.t.1 <- temp(T.1, t)
# Modest climate change:
Temp.t.2 <- temp(T.2, t)
# Extreme change:
Temp.t.3 <- temp(T.3, t)
# Create a data frame:
temp.df <- tibble(t, Temp.t.1, Temp.t.2, Temp.t.3)

# Create a plot of temperature over time for these three situations:
Temp.t.Plot <- 
  # change the df to longer format inorder to use groups:
  temp.df %>% pivot_longer(-t,
                                names_to = "climate_change",
                                values_to = "temperature") %>% 
  ggplot()+
  geom_line(aes(x = t,
                y = temperature,
                color = climate_change))+
  # Add theme and titles and lables:
  theme_bw()+
  labs(title = "Temperature Change (Relative) At Time t Under BAU",
       subtitle = "Three Diffrent Climate Change Scenarios",
       x = "Year t (Relative to time 0)", y = "Temperature Increase")+
  scale_color_discrete(name = "Climate Change Scenario",
                       labels = c("No Climate Change",
                                  "Modest Climate Change",
                                  "Extreme Climate Change"))+
  # Change theme variables:
  theme(text = element_text(family = "serif"),
        plot.title = element_text(size=14, face="bold"),
        plot.subtitle = element_text(size = 9, face = "bold.italic"),
        axis.title.x = element_text(size=10, face="bold", vjust = -1),
        axis.title.y = element_text(size=10, face="bold", vjust = 2),
        strip.text.x = element_text(size = 10, face = "bold.italic"),
        legend.title = element_text(size = 10, face = "bold.italic"),
        legend.position = "bottom")
```

(b) Plot consumption over time for no climate change, modest climate change, and extreme climate change.
```{r}
keep_no_CC <- keep(no_CC)
# = 1
keep_modest_CC <- keep(modest_CC)
# = 0.8187308
keep_extreme_CC <- keep(extreme_CC)
# = 0.0407622

consump_no_CC <- consump(keep_no_CC, 100)
# = 2.718282
consump_modest_CC <- consump(keep_modest_CC, 100)
# = 2.225541
consump_extreme_CC <- consump(keep_extreme_CC, 100)
# = 0.1108032
```


(c) Plot undiscounted utility over time for no climate change, modest climate change, and extreme climate change.
```{r}

util_no_CC <- util(consump_no_CC)
# = 3.297443

util_modest_CC <- util(consump_modest_CC)
# = 2.983649

util_extreme_CC <- util(consump_extreme_CC)
# = 0.6657422
```


2. ANALYSIS

AAAAAAAAAAAAAAAAAAAAAAAAAAAAAA
Suppose T = 4.4
In other words, suppose we know for sure that under BAU, climate change will eventually lead to a 4.4 degree increase in temperature.

What is the present value (i.e. discounted) utility over the next 200 years with climate change?
```{r}
# discount rate = 0.01
t = 200

temp_CC_4.4 <- temp(t = 200, T = 4.4) #4.4
keep_CC_4.4 <- keep(temp_CC_4.4) # 0.379842
consump_CC_4.4 <- consump(keep_CC_4.4, 200) # 2.806674
util_CC_4.4 <- util(consump_CC_4.4) # 3.350626

PV_CC_4.4 = (util_CC_4.4)/((1+r)^200)
# Present Value = 0.4579849

#V0 = (Vt)/((1+r)^t)
# V0 = present value
# Vt = value t years from now
```

What is the present value utility without climate change?
```{r}
temp_CC_no <- temp(t = 200, T = 0) #0
keep_CC_no <- keep(temp_CC_no) #1
consump_CC_no <- consump(keep_CC_no, 200) #7.389056
util_CC_no <- util(consump_CC_no) #5.436564

PV_CC_no = (util_CC_no)/((1+r)^200)
# Present Value = 0.7431042
```

What is the percentage loss in present value utility from climate change (call this L)?
```{r}
# % change = (climate change - no climate change) / no climate change

L <- (((PV_CC_4.4 - PV_CC_no)/(PV_CC_no))*100)
# -38.37 %
```



BBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBBB
Now show how sensitive your calculation of L is to the following parameters: T, g, η, and β. To do so, calculate the % change in L that arises from a 10% increase in each of these parameters.
```{r}
# L = -38.37 %

T_CC_no <- 0
T_CC_4.4 <- 4.4
g <- 0.01
eta <- 0.5
beta <- 0.05

# increase parameters by 10%

T_CC_no_pinc <- T_CC_no*1.1 # = 0
T_CC_4.4_pinc <- T_CC_4.4*1.1 # = 4.84
g_pinc <- g*1.1 # = 0.011
eta_pinc <- eta*1.1 # = 0.55
beta_pinc <- beta*1.1 # = 0.055

# redefine temp fxn
temp_pinc = function(t,T)
{
  out = min(t*T/100,T)
  return(out)
}

# redefine keep fxn
keep_pinc = function(temp_pinc)
{
  out = exp(-beta_pinc*(temp_pinc)^2)
  return(out)
}

# redefine consump fxn
consump_pinc = function(keep_pinc,t)
{
  out = keep_pinc*exp(g_pinc*t)
  return(out)
}

# redefine util fxn
util_pinc = function(C)
{
  out = ((C^(1-eta_pinc))/(1-eta_pinc))
  return(out)
}

# before 10% increase value
#temp_CC_4.4 <- temp(t = 200, T = 4.4) # 4.4
#keep_CC_4.4 <- keep(temp_CC_4.4) # 0.379842
#consump_CC_4.4 <- consump(keep_CC_4.4, 200) # 2.806674
#util_CC_4.4 <- util(consump_CC_4.4) # 3.350626

# recalculate CC of 4.4 w/ pinc
temp_CC_4.4_pinc <- temp(t = 200, T = 4.4)
keep_CC_4.4_pinc <- keep_pinc(temp_CC_4.4_pinc) # 0.3447968
consump_CC_4.4_pinc <- consump_pinc(keep_CC_4.4_pinc, 200) # 3.111796
util_CC_4.4_pinc <- util_pinc(consump_CC_4.4_pinc) # 3.703757

PV_CC_4.4_pinc = (util_CC_4.4_pinc)/((1+r)^200)
# Present Value = 0.5062531

# recalculate no CC w/ pinc
temp_CC_no_pinc <- temp_pinc(t = 200, T = 0) # 0 
keep_CC_no_pinc <- keep_pinc(temp_CC_no_pinc) # 1
consump_CC_no_pinc <- consump_pinc(keep_CC_no_pinc, 200) # 9.025013
util_CC_no_pinc <- util_pinc(consump_CC_no_pinc) # 5.980521


# % change = (climate change - no climate change) / no climate change

L <- (((PV_CC_4.4 - PV_CC_no)/(PV_CC_no))*100)
# -38.37 %

```

