---
title: 'ECON: Homework 4'
author: "Zoe Duerksen-Salm & Renata Massion"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

# Attach packages
library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(effsize)
library(kableExtra)
library(ggbeeswarm)
library(ggridges)
library(patchwork)
library(car)
library(corrplot)
library(stargazer)
library(sf)
library(gganimate)
library(transformr)
```

```{r}
## Define Parameters

delta = 0.005
eta = 0.5
g = 0.01
beta = 0.05

r = delta + eta*g

t <- 0:200

T_0 <- 0
T_2 <- 2
T_4 <- 4
T_4.4 <- 4.4
T_6 <- 6
T_8 <- 8


## Write functions

# Temp
temp = function(t,T)
{
  out = pmin(t*T/100,T)
  return(out)
}

# Keep 
keep = function(beta, temp)
{
  out = exp(-1*beta*temp^2)
  return(out)
}

# Consumption
consump = function(keep,t,g)
{
  out = keep*exp(g*t)
  return(out)
}

# Utility
util = function(C, eta)
{
  out = ((C^(1-eta))/(1-eta))
  return(out)
}
```

---

**Baseline calculations for all questions:**

- Used R Studio for all questions
- We created variables for the given parameters (delta, eta, g, beta, and time t as a sequence of 0-200)
- We then created functions in R Studio for all five given equations (temperature, keep, consumption, utility, and Ramsey's rule).



---

## Question #1: Plots

**Baseline calculations for question #1:**

* Create data frames for no climate change projection/scenario (T=0) and both potential climate change projections/scenarios (T=2 and T=8).
* Include in data frames:
  + **rows** for every year
  + **columns** for results from each function (temperature, keep, consumption, and utility)

---

```{r}
# Make data frames for 3 temps


T_0_df <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(fraction_economy_retained0 = keep(beta,temp)) %>% 
  mutate(consumption0 = consump(fraction_economy_retained0, t, g)) %>% 
  mutate(utility0 = util(consumption0, eta)) %>% 
  mutate(pv_utility0 = utility0/((1+r)^t))

T_2_df <- data.frame(year = t,
                 temp_increase = T_2,
                 temp = temp(t, T_2)) %>% 
  mutate(fraction_economy_retained2 = keep(beta,temp)) %>% 
  mutate(consumption2 = consump(fraction_economy_retained2, t, g)) %>% 
  mutate(utility2 = util(consumption2, eta))

T_8_df <- data.frame(year = t,
                 temp_increase = T_8,
                 temp = temp(t, T_8)) %>% 
  mutate(fraction_economy_retained8 = keep(beta,temp)) %>% 
  mutate(consumption8 = consump(fraction_economy_retained8, t, g)) %>% 
  mutate(utility8 = util(consumption8, eta))
```

#### A. Plot temperature over time

```{r}
ggplot() +
  geom_line(data = T_0_df, aes(x = year, y = temp, color = "No Climate Change"),
            size = 1) +
  geom_line(data = T_2_df, aes(x = year, y = temp, color = "Modest Climate Change"),
            size = 1)+
  geom_line(data = T_8_df, aes(x = year, y = temp, color = "Extreme Climate Change"),
            size = 1)+
  scale_color_manual(values = c("maroon", "orange", "forestgreen"))+
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Increase in Temperature",
       title = "Temperature Over Time",
       color = "Projection") 
```

*Calculations:*

* Create a plot that combines information from the three previously created data frames. Plot year (0 to 200) on the x-axis and increase in temperature (calculated using temperature function/equation) on the y-axis.

---

#### B. Plot consumption over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = consumption0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = consumption2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = consumption8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("maroon", "orange", "forestgreen")) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Consumption",
       title = "Consumption Over Time",
       color = "Projection")
```

*Calculations:*

* Create a plot that combines information from the three previously created data frames. Plot year (0 to 200) on the x-axis and consumption (calculated using consumption function/equation) on the y-axis.

---

#### C. Plot undiscounted utility over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = utility0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = utility2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = utility8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("maroon", "orange", "forestgreen")) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Undiscounted Utility",
       title = "Undiscounted Utility Over Time",
       color = "Projection")
```

*Calculations:*

* Create a plot that combines information from the three previously created data frames. Plot year (0 to 200) on the x-axis and utility (calculated using utility function/equation) on the y-axis.

---

## Question #2: Analysis

---

#### A. T=4.4 under BAU
```{r}

# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

df_4.4_all <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(keep4.4 = keep(beta, temp)) %>% 
  mutate(consump4.4 = consump(keep = keep4.4, t, g)) %>% 
  mutate(util4.4 = util(consump4.4, eta)) %>% 
  mutate(pv_util4.4 = util4.4/((1+r)^t))

df_0_all <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(keep0 = keep(beta, temp)) %>% 
  mutate(consump0 = consump(keep = keep0, t, g)) %>% 
  mutate(util0 = util(consump0, eta)) %>% 
  mutate(pv_util0 = util0/((1+r)^t))

pv_U_cc <- sum(df_4.4_all$pv_util4.4)
pv_U_no_cc <- sum(df_0_all$pv_util0)
# Percent difference
pct_loss_L <- ((pv_U_cc-pv_U_no_cc)/(pv_U_no_cc))*100
```

* Present value utility with climate change: **`r round(pv_U_cc, 2)`**

* Present value utility without climate change: **`r round(pv_U_no_cc, 2)`**

* Percent loss in PV utility from climate change (L): **`r round(pct_loss_L, 2)`%**

*Calculations:*

1. Create two data frames: (a) projections under no climate change and (b) projections under a T=4.4 climate change scenario. 
   + Include rows for each year.
   + Mutate columns to represent each function and the present value of utility.
   + PV equation is [utility/((1+r)^t^)].
2. Individually sum PV columns from each data frame.
3. Calculate percent loss (L)
    + Equation: L = [ (PV with CC - PV without CC) / (PV without CC) ]

---

#### B. Percent change in PV utility (L) that results from a 10% decrease in certain parameters

```{r}
T_new <- 4.4*1.1 #4.84
g_new <- g*1.1 # 0.011
eta_new <- eta*1.1 # 0.55
beta_new <- beta*1.1 # 0.055
delta_new <- delta*1.1 # 0.0055

r_g <- delta + (eta*g_new)
r_eta <- delta + (eta_new*g)
```

```{r}
# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

T_inc <- data.frame(year = t,
                 temp_increase = T_new,
                 temp = temp(t, T_new)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r)^t))

# Sum PV
pv_U_T_new <- sum(T_inc$pv_util)

# New L
pct_loss_L_T_new <- ((pv_U_T_new-pv_U_no_cc)/pv_U_no_cc)*100

# Percent change in L
pct_change_L_T_new <- ((pct_loss_L_T_new -pct_loss_L)/pct_loss_L)*100
```
* Increase in **temperature**: % change in L = **`r round(pct_change_L_T_new, 2)`%**

```{r}
# Increase g by 10%:
# new discount rate

g_inc_4.4 <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g_new)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r_g)^t))

g_inc_0 <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g_new)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r_g)^t))

# Sum PV
pv_U_g_new <- sum(g_inc_4.4$pv_util)
pv_U_no_g <- sum(g_inc_0$pv_util)

# New percent difference (L)
pct_loss_L_g_new <- ((pv_U_g_new - pv_U_no_g)/pv_U_no_g)*100

# Percent change in L
pct_change_L_g_new <- ((pct_loss_L_g_new - pct_loss_L)/pct_loss_L)*100
```
* Increase in **g**: % change in L = **`r round(pct_change_L_g_new, 2)`%**

```{r}
# Increase eta by 10%:

eta_inc_4.4 <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta_new)) %>% 
  mutate(pv_util = utilT/((1+r_eta)^t))

eta_inc_0 <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta_new)) %>% 
  mutate(pv_util = utilT/((1+r_eta)^t))

# Sum PV
pv_U_eta_new <- sum(eta_inc_4.4$pv_util)
pv_U_no_eta <- sum(eta_inc_0$pv_util)

# New percent difference (L)
pct_loss_L_eta_new <- ((pv_U_eta_new - pv_U_no_eta)/pv_U_no_eta)*100

# Percent change in L
pct_change_L_eta_new <- ((pct_loss_L_eta_new - pct_loss_L)/pct_loss_L)*100
```
* Increase in **eta**: % change in L = **`r round(pct_change_L_eta_new, 2)`%**

```{r}
# Increase beta by 10%:

beta_inc_4.4 <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(keepT = keep(beta_new,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r)^t))

beta_inc_0 <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(keepT = keep(beta_new,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r)^t))

# Sum PV
pv_U_beta_new <- sum(beta_inc_4.4$pv_util)
pv_U_no_beta <- sum(beta_inc_0$pv_util)

# New percent difference (L)
pct_loss_L_beta_new <- ((pv_U_beta_new - pv_U_no_beta)/pv_U_no_beta)*100

# Percent change in L
pct_change_L_beta_new <- ((pct_loss_L_beta_new - pct_loss_L)/pct_loss_L)*100
```
* Increase in **beta**: % change in L = **`r round(pct_change_L_beta_new, 2)`%**

*Calculations:*

* Use the same calculation steps from question 2A. BUT, run code 4 separate times (for no climate change scenario and T=4.4 climate change scenario), each run changing increasing one designated variable by 10%.
* **10% change in temp**
  + Increase temperature parameter by 10%. New temp for no CC = 0; new temp for CC 4.4 = 4.84.
    + (old no CC temp * 110%) > (0 * 1.1) > 0
    + (old CC temp * 110%) > (4.4 * 1.1) > 4.84
  + Find the new percent change using the same method as in 2A (summing PVs + % change formula).
  + Use the percent change formula to find the change from the previously calculated L, to the new L [ (New L - Old L) / (Old L) ].
* **10% change in g**
  + Increase g parameter by 10%. New g for both projections = 0.011.
    + (old g * 110%) > (0.01 * 1.1) > 0.011
  + Increasing the g parameter also requires recalculation of the discount value (by plugging the new g into the Ramsey equation).
    + new r = delta + (eta * (new g)) > new r = 0.005 + (0.5 * (0.011)) > new r = 0.0105
  + Find the new percent change using the same method as in 2A (summing PVs + % change formula).
  + Use the percent change formula to find the change from the previously calculated L, to the new L [ (New L - Old L) / (Old L) ].
* **10% change in eta**
  + Increase eta parameter by 10%. New eta = 0.55.
    + (old eta * 110%) > (0.5 * 1.1) > 0.55
  + Increasing the eta parameter also requires recalculation of the discount value (by plugging the new eta into the Ramsey equation).
    + new r = delta + ((new eta) * g) > new r =  0.005 + ((0.55) * 0.01) > new r = 0.0105
  + Find the new percent change using the same method as in 2A (summing PVs + % change formula).
  + Use the percent change formula to find the change from the previously calculated L, to the new L [ (New L - Old L) / (Old L) ].
* **10% change in beta**
  + Increase beta parameter by 10%. New beta = 0.055.
    + (old beta * 110%) > (0.05 * 1.1) > 0.055
  + Find the new percent change using the same method as in 2A (summing PVs + % change formula).
  + Use the percent change formula to find the change from the previously calculated L, to the new L [ (New L - Old L) / (Old L) ].

---

#### C. Maximum value of θ society would be willing to endure every year to completely prevent climate change.
```{r}
#theta^(1-eta) * npv_pv_u_4.4 = npv_pv_u_0

# Consumption
consump_theta = function(keep,t,g, theta)
{
  out = (keep*exp(g*t))*(1-theta)
  return(out)
}

theta <- function(pv_U_cc, pv_U_no_cc)
{
  output <- 1 - (pv_U_cc/pv_U_no_cc)^((1-eta)^-1)
  return(output)
}
theta_star <- (theta(pv_U_cc,pv_U_no_cc))*100

# 1-(pv_U_cc/pv_U_no_cc)^(1/(1-eta))
```
θ∗ = **`r round(theta_star, 2)`%**

*Calculations:*

1. Set PV of discounted utility when T=4.4 equal to PV of discounted utility when T=0 multiplied by theta.
    + PV(4.4) = PV(0)*(theta)
2. Find NPV for each climate change scenario (4.4 vs. 0).
    + NPV(4.4) = NPV(0)*theta^(1-eta)^
3. Rearrange equation such that theta is on the left side of the equation.
    + theta = (NPV(4.4)/NPV(0))^(1/(1-eta))^
4. Plug NPV values into the equation.
    + theta = (198.66/255.2734)^(1/(1-0.5))^
    + theta = 0.6056333
5. Calculate theta* by doing 1-theta found in step 4 and multiplying by 100 to get a percentage.
    + theta* = (1 - 0.6056333)*100
    + theta* = 39.44%

---

#### D. θ∗ under uncertainty over T.
```{r}
# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

df_uncertain <- data.frame("year"= t) %>%
  mutate(temp2 = temp(t, T_2),
         keep2 = keep(beta, temp2),
         consump2 = consump(keep2, t, g),
         util2 = util(consump2, eta),
         pv_u_2 = util2/((1+r)^t),
         temp4 = temp(t, T_4),
         keep4 = keep(beta, temp4),
         consump4 = consump(keep4, t, g),
         util4 = util(consump4, eta),
         pv_u_4 = util4/((1+r)^t),
         temp6 = temp(t, 6),
         keep6 = keep(beta, temp6),
         consump6 = consump(keep6, t, g),
         util6 = util(consump6, eta),
         pv_u_6 = util6/((1+r)^t))
         
npv_pv_u_2 <- sum(df_uncertain$pv_u_2)
npv_pv_u_4 <- sum(df_uncertain$pv_u_4)
npv_pv_u_6 <- sum(df_uncertain$pv_u_6)

npv_pv_u_uncertain <- (0.2 * npv_pv_u_2) + (0.5 * npv_pv_u_4) + (0.3 * npv_pv_u_6)

theta_uncertain <- (theta(npv_pv_u_uncertain, pv_U_no_cc))*100

# 1-(npv_pv_u_uncertain/pv_U_no_cc)^(1/(1-eta))
```
Uncertain θ = **`r round(theta_uncertain, 2)`%**

*Calculations:*



1. Create a data frame for projections under 3 climate change scenarios (T=2, T=4, and T=6) projections under a T=4.4 climate change scenario. 
   + Include rows for each year.
   + Mutate columns to represent calculated answer for each function and present value for a given projection.
2. Find NPV for each scenario by summing columns for PV values.
3. Determine total NPV under uncertainty using the equation: NPV = (p(2) * PV(2)) + (p(4) * PV(4)) + (p(6) * PV(6))
    + Where p = probability of that scenario.
    + NPV = (0.2 * 241.597) + (0.5 * 206.907) + (0.3 * 165.2395)
    + NPV = 201.3447
4. Plug NPV values for the uncertain scenario and the into the T=0 scenario into the equation in step 3 of question 2C.
    + NPV(uncertain) = NPV(0)*theta^(1-eta)^
5. Rearrange equation such that theta is on the left side of the equation.
    + theta = (NPV(uncertain)/NPV(0))^(1/(1-eta))^
6. Plug NPV values into the equation.
    + theta = (201.3447/255.2734)^(1/(1-0.5))^
    + theta = 0.622113
7. Calculate theta* by doing 1-theta found in step 6 and multiplying by 100 to get a percentage.
    + theta* = (1 - 0.622113)*100
    + theta* = 37.79%