---
title: 'ECON: Homework 4'
author: "Zoe Duerksen-Salm & Renata Massion"
date: "5/25/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

# Attach packages
library(tidyverse)
library(tidyr)
library(janitor)
library(here)
library(effsize)
library(kableExtra)
library(ggbeeswarm)
library(ggridges)
library(patchwork)
library(car)
library(corrplot)
library(stargazer)
library(sf)
library(gganimate)
library(transformr)
```

```{r}
## Define Parameters

delta = 0.005
eta = 0.5
g = 0.01
beta = 0.05

r = delta + eta*g

t <- 0:200

T_0 <- 0
T_2 <- 2
T_4 <- 4
T_4.4 <- 4.4
T_6 <- 6
T_8 <- 8


## Write functions

# Temp
temp = function(t,T)
{
  out = pmin(t*T/100,T)
  return(out)
}

# Keep 
keep = function(beta, temp)
{
  out = exp(-1*beta*temp^2)
  return(out)
}

# Consumption
consump = function(keep,t,g)
{
  out = keep*exp(g*t)
  return(out)
}

# Utility
util = function(C, eta)
{
  out = ((C^(1-eta))/(1-eta))
  return(out)
}
```
---

---

## Question #1: Plots
```{r}
# Make data frames for 3 temps


T_0_df <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(fraction_economy_retained0 = keep(beta,temp)) %>% 
  mutate(consumption0 = consump(fraction_economy_retained0, t, g)) %>% 
  mutate(utility0 = util(consumption0, eta)) %>% 
  mutate(pv_utility0 = utility0/((1+r)^t))

T_2_df <- data.frame(year = t,
                 temp_increase = T_2,
                 temp = temp(t, T_2)) %>% 
  mutate(fraction_economy_retained2 = keep(beta,temp)) %>% 
  mutate(consumption2 = consump(fraction_economy_retained2, t, g)) %>% 
  mutate(utility2 = util(consumption2, eta))

T_8_df <- data.frame(year = t,
                 temp_increase = T_8,
                 temp = temp(t, T_8)) %>% 
  mutate(fraction_economy_retained8 = keep(beta,temp)) %>% 
  mutate(consumption8 = consump(fraction_economy_retained8, t, g)) %>% 
  mutate(utility8 = util(consumption8, eta))
```

#### A. Plot temperature over time

```{r}
ggplot() +
  geom_line(data = T_0_df, aes(x = year, y = temp, color = "No Climate Change"),
            size = 1) +
  geom_line(data = T_2_df, aes(x = year, y = temp, color = "Modest Climate Change"),
            size = 1)+
  geom_line(data = T_8_df, aes(x = year, y = temp, color = "Extreme Climate Change"),
            size = 1)+
  scale_color_manual(values = c("maroon", "orange", "forestgreen"))+
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Increase in Temperature",
       title = "Temperature Over Time",
       color = "Projection") 
```

#### B. Plot consumption over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = consumption0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = consumption2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = consumption8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("maroon", "orange", "forestgreen")) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Consumption",
       title = "Consumption Over Time",
       color = "Projection")
```

#### C. Plot undiscounted utility over time

```{r}
ggplot() +
  geom_line(data = T_0_df,
            aes(x = year,
                y = utility0,
                color = "No climate change"),
            size = 1)+
  geom_line(data = T_2_df,
            aes(x = year,
                y = utility2,
                color = "Modest climate change"),
            size = 1)+
  geom_line(data = T_8_df,
            aes(x = year,
                y = utility8,
                color = "Extreme climate change"),
            size = 1) +
  scale_color_manual(values = c("maroon", "orange", "forestgreen")) +
  theme_light() +
  theme(legend.position = "bottom") +
  labs(x = "Year",
       y = "Undiscounted Utility",
       title = "Undiscounted Utility Over Time",
       color = "Projection")
```

---

---

## Question #2: Analysis

#### A. T=4.4 under BAU
```{r}

# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

df_4.4_all <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(keep4.4 = keep(beta, temp)) %>% 
  mutate(consump4.4 = consump(keep = keep4.4, t, g)) %>% 
  mutate(util4.4 = util(consump4.4, eta)) %>% 
  mutate(pv_util4.4 = util4.4/((1+r)^t))

df_0_all <- data.frame(year = t,
                 temp_increase = T_0,
                 temp = temp(t, T_0)) %>% 
  mutate(keep0 = keep(beta, temp)) %>% 
  mutate(consump0 = consump(keep = keep0, t, g)) %>% 
  mutate(util0 = util(consump0, eta)) %>% 
  mutate(pv_util0 = util0/((1+r)^t))

pv_U_cc <- sum(df_4.4_all$pv_util4.4)
pv_U_no_cc <- sum(df_0_all$pv_util0)
# Percent difference
pct_loss_L <- ((pv_U_cc-pv_U_no_cc)/(pv_U_no_cc))*100
```

Present value utility with climate change: **`r round(pv_U_cc, 2)`**

Present value utility without climate change: **`r round(pv_U_no_cc, 2)`**

Percent loss in PV utility from climate change (L): **`r round(pct_loss_L, 2)`%**

---

#### B. Percent change in PV utility (L) that results from a 10% decrease in certain parameters

```{r}
T_new <- 4.4*1.1 #4.84
g_new <- g*1.1 # 0.011
eta_new <- eta*1.1 # 0.55
beta_new <- beta*1.1 # 0.055
delta_new <- delta*1.1 # 0.0055

r_g <- delta + (eta*g_new)
r_eta <- delta + (eta_new*g)
```

##### Increase in temperature
```{r}
# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

T_inc <- data.frame(year = t,
                 temp_increase = T_new,
                 temp = temp(t, T_new)) %>% 
  mutate(keepT = keep(beta,temp)) %>% 
  mutate(consumpT = consump(keepT, t, g)) %>% 
  mutate(utilT = util(consumpT, eta)) %>% 
  mutate(pv_util = utilT/((1+r)^t))

pv_U_T_new <- sum(T_inc$pv_util)

# New L
pct_loss_L_T_new <- (pv_U_T_new-pv_U_no_cc)/pv_U_no_cc

# Percent change in L
pct_change_L_T_new <- (pct_loss_L_T_new -pct_loss_L)/pct_loss_L
```

##### Increase in g
```{r}
# Increase g by 10%:
# new discount rate

g_inc <- data.frame(year = t,
                 temp_increase = T_4.4,
                 temp = temp(t, T_4.4)) %>% 
  mutate(fraction_economy_retained = econ(beta,temp)) %>% 
  mutate(consumption = C(fraction_economy_retained, g_new, year)) %>%
  mutate(utility = U(consumption, n)) %>% 
  mutate(pv_utility = utility/((1+r)^t))
# Sum PV
pv_U_g_new <- sum(g_new_df$pv_utility)
# New percent difference (L)
pct_loss_L_g_new <- (pv_U_g_new-pv_U_without_cc)/pv_U_without_cc
# Percent change in L
pct_change_L_g_new <- (pct_loss_L_g_new -pct_loss_L)/pct_loss_L
```

##### Increase in eta
```{r}

```

##### Increase in beta
```{r}

```

---

#### C. Maximum value of θ society would be willing to endure every year to completely prevent climate change.
```{r}
#theta^(1-eta) * npv_pv_u_4.4 = npv_pv_u_0
theta <- function(pv_U_cc, pv_U_no_cc)
{
  output <- 1 - (pv_U_cc/pv_U_no_cc)^((1-eta)^-1)
  return(output)
}
theta_star <- theta(pv_U_cc,pv_U_no_cc)
```
θ∗ = **`r round(theta_star, 2)`**

---

#### D. θ∗ under uncertainty over T.
```{r}
# temp(t, T)
# keep(beta, temp)
# consump(keep, t, g)
# util(C, eta)

df_uncertain <- data.frame("year"= t) %>%
  mutate(temp2 = temp(t, T_2),
         keep2 = keep(beta, temp2),
         consump2 = consump(keep2, t, g),
         util2 = util(consump2, eta),
         pv_u_2 = util2/((1+r)^t),
         temp4 = temp(t, T_4),
         keep4 = keep(beta, temp4),
         consump4 = consump(keep4, t, g),
         util4 = util(consump4, eta),
         pv_u_4 = util4/((1+r)^t),
         temp6 = temp(t, 6),
         keep6 = keep(beta, temp6),
         consump6 = consump(keep6, t, g),
         util6 = util(consump6, eta),
         pv_u_6 = util6/((1+r)^t))
         
npv_pv_u_2 <- sum(df_uncertain$pv_u_2)
npv_pv_u_4 <- sum(df_uncertain$pv_u_4)
npv_pv_u_6 <- sum(df_uncertain$pv_u_6)

npv_pv_u_uncertain <- (0.2 * npv_pv_u_2) + (0.5 * npv_pv_u_4) + (0.3 * npv_pv_u_6)

theta_uncertain <- theta(npv_pv_u_uncertain, pv_U_no_cc)
```
Uncertain θ = **`r round(theta_uncertain, 2)`**
